A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway

### 2018-01-01 TUTORIAL.Rmd
edited tutorial, made manuscript skeleton

---
title: "Collaborative analysis and writing in R"
author: "R Clay Wright"
date: "12/28/2017"
output: html_document
---

@@

This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: " natural variation in the nuclear auxin response"). I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio, 
but if you feel you could use a refresher check out [this introduction](https://github.com/wrightrc/Intro_to_R_RStudio_and_ANOVA). We'll use docker to share with you a version of R and Rstudio that already has all of the necessary packages to do the analysis. We'll use Rmarkdown to write the manuscript and make figures and tables by integrating prose, R code and results from the code.

So in this tutorial I will guide you through, in as abstracted a way as possible:  
1. Installing docker  
2. Starting the version of RStudio in docker  
3. Working in docker/RStudio  
4. Saving your work in docker/RStudio  
5. Using Rmarkdown to format text  
6. Using Rmarkdown to insert R code and results into text  
7. Citations in Rmarkdown

@@

## docker

@@

>An **image** is a lightweight, stand-alone, executable package that includes everything needed to run a piece of software, including the code, a runtime, libraries, environment variables, and config files.  
>  
>A container is a runtime instance of an image—what the image becomes in memory when actually executed. It runs completely isolated from the host environment by default, only accessing host files and ports if configured to do so.

@@

The image that we are going to work with here is called `r1001genomes`. Docker images build from previous images (like all programming does, and science for that matter). I built the `r1001genomes` image on top of a previous image called `verse` from the `rocker` project. The `verse` image contains R, Rstudio, the [tidyverse](www.tidyverse.org) packages, Rmarkdown and all of the dependencies for these things. To this, I added  the necessary packages that `r1001genomes` builds from and `r1001genomes`. Let's make a new container for this image!

@@

Kitematic will then restart the container and the access URL might be changed. So go back to the "Home" tab and paste the new URL into your browser or, if the access URL didn't change, reload the page.

Now when you click on the "kitematic" folder you will see the contents of this folder. Set the working directory to the "manuscript" folder (using code, or the "More" button). Click on "manuscript.Rmd" to open the manuscript markdown file.

@@

The idea behind markdown and R markdown was to make html (hyper-text markUP language) more low key. So the goal of markDOWN, is to be easy to write and easy to read but still have a lot of the powers of HTML (text formatting, layout and templating).

This document was is actually written in R Markdown. If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane.  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.45.46.png)

@@

### Including Plots

@@

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Don't get bogged down in formatting the text, but I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)! You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu. Focus foremost should be on the content. We'll worry about formatting in revisions.
![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.57.15.png)

In the manuscript skeleton and the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document.

### Citations in R markdown  
You can include citations in markdown documents using the following syntax: `[@citation_key]`. The citation key refers to the unique identifier generated when exporting a bibliography from a reference manager. For the manuscript you will need to export the references from your reference editor in one of many formats, but I would recommend BibTex (.bib). For other formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).

Let's take a look at an example, open the "references.bib" file in the "manuscript folder". You'll notice each reference begins with an @-symbol and the type of reference. Then within curly braces are all of the data about the reference. The first item is the citation key, which in this case is "gray_identification_1999". So to cite this paper in markdown you would add `[@gray_identification_1999]` to the end of a sentence. My process for citing-while-writing is to first write, then when I want to cite something, find the paper in my reference manager and copy it to a folder for the project I'm working on. This will allow me to export the folder as a .bib file later). Then I copy the citation key for the reference and paste it into my writing. When I'm ready to knit the document, I'll export the folder of references as a .bib (with Better Bibtex you can sync this file so that it is automatically updated). This bibliography file must be in the same folder with the .Rmd file, along with a .csl file that specifies the format of the citations and bibliography in the final document (available for most journals [via Zotero](https://www.zotero.org/styles)). These files also must be specified in the header of the document.  
![](TUTORIAL_screenshots/Screenshot 2017-12-30 14.04.14.png)

Most common reference managers allow you to export a bibliography as a bibtex file, and copy and paste citation keys, however I would highly recommend downloading [Zotero](www.zotero.org) and importing your reference library from whatever reference editor you currently use. Zotero is free, open-source and integrates very well with all popular web browsers and writing tools. If you want to work in a different citation manager, here are some links to help you create and copy and paste citation keys as well as export BibTex bibliography files. Note that most of these tutorials are for Latex, which uses a slightly different syntax from Rmarkdown. In Rmarkdown insert citations via `[@citation_key]`. The `citation_key` string does have some stipulations 
>The citation key must begin with a letter, digit, or \_, and may contain alphanumerics, _, and internal punctuation characters (:.#$%&-+?<>~/).

@@

>>>>

---
title: "Collaborative analysis and writing in R"
author: "R Clay Wright"
date: "12/28/2017"
output: 
  html_document:
    toc: true

---

@@

This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: "natural variation in the nuclear auxin response"). I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio, 
but if you feel you could use a refresher check out [this introduction](https://github.com/wrightrc/Intro_to_R_RStudio_and_ANOVA). We'll use docker to share with you a version of R and Rstudio that already has all of the necessary packages to do the analysis. We'll use Rmarkdown to write the manuscript and make figures and tables by integrating prose, R code and results from the code.

So in this tutorial I will guide you through:  
1. Installing docker  
2. Starting RStudio and r1001genomes in docker  
3. Working in docker/RStudio  
4. Saving your work in docker/RStudio  
5. Using Rmarkdown to format text  
6. Using Rmarkdown to insert R code and results into text  
7. Citations in Rmarkdown

@@

## Docker

@@

>An **image** is a lightweight, stand-alone, executable package that includes everything needed to run a piece of software, including the code, a runtime, libraries, environment variables, and config files.  
>  
>A **container** is a runtime instance of an image—what the image becomes in memory when actually executed. It runs completely isolated from the host environment by default, only accessing host files and ports if configured to do so.

@@

The image that we are going to work with here is called `r1001genomes`. Docker images build on previous images (like all programming does, and all science does, for that matter). I built the `r1001genomes` image on top of a previous image called `verse` from the `rocker` project. (The rocker project and R Open Science also have a [great docker tutorial](http://ropenscilabs.github.io/r-docker-tutorial/).) The `verse` image contains R, Rstudio, the [tidyverse](www.tidyverse.org) packages, Rmarkdown and all of the dependencies for these things. To this, I added  the necessary packages that `r1001genomes` builds from and `r1001genomes`. Let's make a new container for this image!

@@

Kitematic will then restart the container and the access URL might be changed. So go back to the "Home" tab and paste the new URL into your browser or, if the access URL didn't change just reload the page.

Now when you click on the "kitematic" folder you will see the contents of this folder. Any files you save in this folder will be saved to the folder you mapped on your computer. You can also open files from this folder in RStudio. In the RStudio "Files" pane, navigate to the "manuscript" folder and set it as your working directory (using the "More" button). Click on "manuscript.Rmd" to open the manuscript markdown file.

@@

The idea behind the creation of markdown and R markdown was to make html (hyper-text markUP language) more low key. So the goals of markDOWN are to be easy to write and easy to read but to still have a lot of the powers of HTML (text formatting, layout and templating). R Markdown also allows us to generate PDF, Word, and Presentation documents in addition to HTML.

This document was is actually written in R Markdown. If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane (see the figure).  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.45.46.png)

@@

You can insert code chunks easily using the "Insert" button at the top of the source window, Code>Insert Chunk in the menu bar or hotkeys specified next to this menu bar command.

### Including Plots and Code

@@

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot (check out the markdown code in TUTORIAL.Rmd).

You can also incorporate results from code chunks in your prose. First, we need a result .

```{r}
set.seed(42)
group1 <- rnorm(n = 10, mean = 1, sd = 0.5)
group2 <- rnorm(n = 10, mean = 2, sd = 0.5)
t_test <- t.test(x = group1, y = group2)
t_test
```

Now we can refer to these results in our code by creating an inline code chunk with syntax `r #code`. So we could say that we have some support that the means of the two groups we simulated are different (`r t_test$method`, t-value = `r t_test$statistic`, p = `r t_test$p.value`)

Don't get bogged down in formatting the text, but I would highly recommend checking out the <http://rmarkdown.rstudio.com> link and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)! You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu.  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.57.15.png)

A few tips to help your R Markdown experience stay low key:  
1. Turn on whitespace characters so you know where spaces, tabs and returns are. This is in Preferences>Code>Display.  
2. Include 2 spaces at the end of each paragraph or whenever you want to start a new line (like in this numbered list).  
3. Use and empty line between paragraphs, headings, offset quotes and any other sort of formatted elements like lists, tables and code chunks. This offset helps improve readability and makes sure that these elements are formatted the way you want them.

**However, keep in mind your foremost focus should be on the content. We'll worry about formatting in revisions.**

In the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document. But, one last thing before you move on to the example analysis and your own analysis...

### Citations in R markdown

You can include citations in markdown documents using the following syntax: `[@citation_key]`. The citation key refers to a unique identifier generated when exporting a bibliography from a reference manager. For the manuscript you will need to export the references from your reference editor in one of many formats, but I would recommend BibTex (.bib). For other possible formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).

Let's take a look at an example, open the "references.bib" file. You'll notice each reference begins with an @-symbol and the type of reference. Then within curly braces are all of the data about the reference. The first item is the citation key, which in this case is "gray_identification_1999". So to cite this paper in markdown you would add `[@gray_identification_1999]` to the end of a sentence. My process for citing-while-writing is to first write, then when I want to cite something, find the paper in my reference manager and copy it to a folder for the project I'm working on. This will allow me to export the folder as a .bib file later). Then I copy the citation key for the reference and paste it into my writing. When I'm ready to knit the document, I'll export the folder of references as a .bib (with Better Bibtex you can sync this file so that it is automatically updated). This bibliography file must be in the same folder with the .Rmd file, along with a .csl file that specifies the format of the citations and bibliography in the final document (available for most journals [via Zotero](https://www.zotero.org/styles)). These files also must be specified in the header of the document.  
![](TUTORIAL_screenshots/Screenshot 2017-12-30 14.04.14.png)

Most common reference managers allow you to export a bibliography as a bibtex file, and copy and paste citation keys, however I would highly recommend downloading [Zotero](www.zotero.org) and importing your reference library from whatever reference editor you currently use. Zotero is free, open-source and integrates very well with all popular web browsers and writing tools. If you want to work in a different citation manager, here are some links to help you create and copy and paste citation keys as well as export BibTex bibliography files. Note that most of these tutorials are for Latex, which uses a slightly different syntax from Rmarkdown. In Rmarkdown insert citations via `[@citation_key]`. The `citation_key` string does have some stipulations,

>The citation key must begin with a letter, digit, or \_, and may contain alphanumerics, _, and internal punctuation characters (:.#$%&-+?<>~/).

This can cause some issues with BibTex files exported from some reference managers, that is why I recommend Zotero with the Better BibTex add-on.

@@

Now you are ready to work through example.Rmd (and/or example.html) and begin thinking about and analyzing natural variation in your gene family!

### 2018-01-01 manuscript/manuscript.Rmd
edited tutorial, made manuscript skeleton

---
title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
author: 
  affiliation: University of Washington
  name: "R. Clay Wright"
author:

@@

output:
  html_document:
     fig_caption: yes
#  word_document:
#    fig_caption: yes
#    fig_height: 6
#    fig_width: 6
#    highlight: haddock
keywords: r markdown
fontsize: 11pt
fontfamily: Helvetica
output: 
   pdf_document: 
     toc: true
     number_sections: true
bibliography: ["references.bib"]
csl: apa.csl
link-citations: true
---

@@

#List of working titles  
You should keep a list of titles and constantly be refining and tweaking it. A title helps focus you. The title should be in the same style as papers you plan to cite.
You might also think of snappy titles or sub titles for press releases. Find some press releases for paper similar to yours to see how journalists interpret the results.

#Abstract  
Your abstract should be in the following form, more or less exactly:

* Motivating, high level statement  
* Statement defining the cutting edge of your research area  
* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
addressing.  
* A statement of the form:"Here, we ..."  that concisely states your contribution.  
* A series of 3?5 statement about your results.  
* A concluding statement.  
When presenting your abstract to your advisor, these sentences should be clearly laid out, in a bullet list preferably.

#Related work  
You need a list of about 30 papers that are the most relevant. They should be weighted toward recent, well cited work. You should group them into two to four areas. For each paper, you should have a sentence that says what the paper is about and how it relates to your work.

#Glossary  
Make a list of all of the new terms that you have invented to talk about different parts of your work. This list will eventually be put into paragraph format, but for now it should look like a glossary. You can also add terms that from the literature that you will use in case you want to make sure that the reader (or you or your advisor) really knows what you mean when you use that term.

#Results  
This should be a list of 3-5  statements  that very exactly say what you are claiming as results, and then the support for that claim. Statements will be then supported by experimental, theoretical, or numerical evidence.

Note that it is very important that you do not use mere titles for the sections of your paper. If you do, you end up with a laundry list. So do not say "transcription initiation" and list all of the things you know about the subject. Instead say, say "Transcription initiation is regulated by proteins that bind to or near promoters", and then list the evidence you have (or plan to get) that says why you suspect the sentence to be true.

Another way to think of these statements is that they are essentially your hypotheses, if you are a scientist, or they are statements of capabilities if you are an engineer.

#Figures and captions

You should have mock ups of three to five figures (with sub?figures) that you plan to get the data for. These figures are incredibly important to think about from an early stage. For example, if you know that you will have a certain data figure with certain x and y axes and certain error bars, then you have a specification of what data you will need to generate and what it ought to look like.

You should also have working drafts of all of the captions for each sub figure. Captions have  titles and then descriptions.

Finally, you should know what glossary terms or results statements the figures correspond to and roughly how you will use them in your paper.

#List of things to do

Keep a running list of everything you need to do to finish the paper. Don’t put anything extraneous on the list. Make sure every unfinished result has a corresponding sub?list of things to do. Make sure that there are always items on your list that you can and will do t  his week,  as well as longer term items that help you plan ahead.

@@

>>>>

---
title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
author: 
  affiliation: University of Washington
  name: "R. Clay Wright"

@@

output: 
    html_document:
      toc: true
      fig_caption: yes

keywords: r markdown
fontsize: 11pt
fontfamily: Helvetica

bibliography: ["natural_variation.bib", "TIR1_AFB.bib"]
csl: apa.csl
link-citations: true
---

@@

# List of working titles  
* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
* *Add your suggestions*

@@

* Motivating, high level statement  
* Statement defining the cutting edge of your research area  
* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
addressing.  
* A statement of the form:"Here, we ..."  that concisely states your contribution.  
* A series of 3?5 statement about your results.  
* A concluding statement.

# Related work  
What and how have we, as a field, learned from natural variation?  
[@hollister_genomic_2014; @bielawski_detecting_2013; @kryazhimskiy_population_2008; @schmid_multilocus_2005; @nieduszynski_sequence_2011; @melamed_combining_2015; @weigel_population_2015; @schmitz_patterns_2013; @weigel_natural_2005; @hardtke_case_2010; @koornneef_naturally_2004; @mitchell-olds_genetic_2006; @joly-lopez_developing_2016; @wright_insights_2017a; @park_distributions_2017; @_135_2016a]

# Methods

# Results  
## TIR1/AFB genes  
[@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]

## Aux/IAA genes

### Group A Aux/IAAs

### Group B Aux/IAAs

### Group C Aux/IAAs

## TPL/TPR genes

## ARF genes

### Class A ARFs

### Class B ARFs

### Class C ARFs
